### 2024-12-03 paper/paper.qmd
Began to add in data section content, need to still make tables within this section

---
title: "..."
subtitle: "..."
author: 
  - Julia Lee
thanks: "Code and data are available at: https://github.com/jjlee-lee/Toronto_Neighbourhood_Income.git"
date: today
date-format: long
abstract: "..."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(tinytex)
library(knitr)
library(ggplot2)
library(dplyr)
library(arrow)
library(here)
library(modelsummary)
```

# Introduction

# Data {#sec-data}

To simulate, test, download, and clean the Neighbourhoods Profile data, the statistical programming language R was used [@citeR]. Specific libraries that assisted the analysis include `tidyverse` [@tidyverse], `dplyr` [@dplyr], `opendatatoronto` [@opendatatoronto], `tinytex` [@tinytex], `ggplot2` [@ggplot2], `knitr` [@knitr], `testthat` [@testthat], `here` [@here], `arrow` [@arrow], and `modelsummary` [@modelsummary].

## Neighbourhood Profiles Data

Neighbourhood Profiles data for the city of Toronto is provided by the Social Development, Finance & Administration from Toronto’s Open Data portal [@neighbourhoodprofile]. The dataset obtained from the city of Toronto’s Open Data portal [@opentorontodata], is a collection of data from the 2021 Canadian Census from Statistics Canada [@statisticscanada]. Records of socio-economic data for 158 geographic regions (i.e. social planning neighbourhoods) in Toronto can be found within this dataset. From age to ethnocultural diversity, the Census data highlights both the socio-economic and demographic characteristics of Toronto residents across its individual neighbourhoods.

The Neighbourhood Profiles data is an extensive dataset with each row representing a demographic or socio-economic characteristic and each column reflecting a Toronto neighbourhood. For each neighbourhood, there is information on its name, identification number, and status. Status refers to a neighbourhood’s Toronto Strong Neighbourhoods Strategy (TSNS) designation, which indicates whether a neighbourhood is an emerging area, an area that needs improvement, or neither.  Additionally, data for each neighbourhood’s age population, income, education, and more from the 2021 Census is provided. @tbl-rawdata below offers a preview of this dataset.

@tbl-rawdata displays neighbourhood names, numbers, and TSNS designations along with data for the number of individuals within different age groups in each neighbourhood. For example, Thistletown-Beaumond Heights is a neighbourhood in Toronto that needs improvement and has a fairly large number of children ages 0 to 14 years old. The rest of the dataset follows the same format, displaying different characteristics of all 158 neighbourhoods in the city of Toronto.

## Analysis Data

For the present analysis, the variables of interest are the different types of census family sizes along with the average after-tax income for each of Toronto’s neighbourhoods. The Neighbourhood Profiles data includes four different census family sizes: (1) 2-person families, (2) 3-person families, (3) 4-person families, and (4) five or more-person families. It also provides the average after-tax income ($) recorded in 2020 for each neighbourhood.

Since the objective is to find out what kinds of families are driving average neighbourhood income in Toronto, the data used throughout this analysis reflects the average income level for all 158 neighbourhoods alongside their counts of different census family sizes. With minimal data wrangling, the analysis data is simply an extraction of the larger dataset with a focus on the different family sizes and average after-tax income. @tbl-analysis below illustrates this analysis data for three neighbourhoods and summary statistics for this data can be found in the Appendix (@sec-appendix).

```{r}
#| echo: false
#| message: false
#| label: tbl-analysis
#| tbl-cap: "Toronto Neighbourhood Profiles Data"

analysis <- read_parquet(here::here("data/02-analysis_data/analysis_data.parquet"))
#view(analysis)

for_table <- analysis[,c(1,2,4,5,6,7,10)]

for_table |>
  slice(1:3) |>
  kable(
    col.names = c("Name", "Number", "2 persons", "3 persons",
                  "4 persons", "5 or more persons", "Average Income"),
    digits = 1,
    linesep = "",
    align = c("l", "l", "c", "c", "c", "c", "c")
  )
```

In @tbl-analysis, the “Name” column represents each neighbourhood’s name, and the “Number” column reflects each neighbourhood’s identification number. The variables, “2 persons”, “3 persons”, “4 persons”, and “5 or more persons” represent the number of census families in each neighbourhood with those particular family sizes. Lastly, the “Average Income” variable reflects the average after-tax income for all neighbourhoods in 2020 (recorded in dollars). Within the analysis data file provided by this analysis, there are additional variables that represent the total number of census families by family size, the average census family size, and the average number of children in census families for each neighbourhood. These variables are considered to obtain additional context about how families are made up and distributed across Toronto’s neighbourhoods. It is important to note that while this additional information provided a better understanding of the variables of interest, they are not included in the model of the analysis. A detailed account of the model can be found in (@sec-model).

## Map Data

To further understand the type of families that drive average neighbourhood income, this analysis also uses Neighbourhoods data – a shapefile that contains geographic information of Toronto neighbourhood boundaries – to map the results of its model. The boundaries of each neighbourhood are defined using census tract information provided by Statistics Canada, and the shapefile itself is published by the Social Development, Finance & Administration from Toronto’s Open Data portal [@socialdevelopment]. The map within this analysis is created using ArcGIS Pro software [@arcgis]. By joining the analysis data shown in @tbl-analysis with the shapefile based on neighbourhood names, a map that highlights the distribution of family sizes that drive average neighbourhood income is built. The information contained in the shapefile is shown below in @tbl-mapdata.

The Neighbourhoods shapefile data contains neighbourhood names, numbers, and TSNS designations similar to the Neighbourhood Profiles data. It also includes geographic information (i.e. Shape_Length and Shape_Area) that defines the boundaries of each neighbourhood.

## A Note on Measurement

The Neighbourhood Profiles data is a reflection of the 2021 Census for neighbourhoods in Toronto, meaning that it measures and represents the demographic and socio-economic characteristics of families across Toronto at this particular time period. This information is collected through the use of questionnaires – a short and long-form census. The short-form census is sent out to all households, and it asks questions about simple characteristics like age, gender, and household size. With this information, the short-form census attempts to enumerate all individuals within a geographic region and may impute any missing data. The long-form census is sent out to only 25% of households and is a deeper source of data that outlines information such as housing, education, and ethnicity.

So, various aspects of families’ livelihoods are measured through the Census and aggregated to create an overall picture of particular geographic regions – in this case, Toronto neighbourhoods. It is important to note that through this measurement process, there is room for error as well as a need to be aware of this potential for error. Missing data values can be imputed and the aggregation of individual household characteristics or experiences can lead to incorrect assumptions about certain regions and their populations. Thus, this analysis presents its findings with this in mind.

# Model

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

## Model justification

## Model Validation

# Results

# Discussion

## First discussion point {#sec-first-point}

## Second discussion point

## Third discussion point

## Limiations & Future Directions

\newpage

\appendix

# Appendix {#sec-appendix}

## Analysis Data Summary Statistics

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References