### 2024-12-04 plan/datasheet/datasheet_template.qmd
Updated datasheet for analysis data

---
title: "Datasheet for Toronto Neighbourhood Profiles Data"
subtitle: "A deeper look into the 2021 Canadian Census Data"
author: 
  - Julia Lee
thanks: "Code and data are available at: https://github.com/jjlee-lee/Toronto_Neighbourhood_Income.git"
date: today
date-format: long
abstract: "The following contains information about the Toronto Neighbourhoods Profile Data. Extract of the questions are provided by Gebru and colleagues (2021, https://doi.org/10.1145/3458723)"
format: pdf
number-sections: true
---

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The Neighbourhood Profiles dataset was created to gain a better understanding of the populations living in the city of Toronto. By highlighting various demographic and socio-economic characteristics such as education, income, and employment of Toronto residents, this data can be used to understand how Toronto is changing and to better serve all Toronto residents. 
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - This dataset is a collection of data from the 2021 Canadian Census conducted and provided by Statistics Canada. To organize population information for Toronto specifically, the Social Development, Finance & Administration of Open Data Toronto created this dataset. 
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - The creation of the dataset is funded by the city of Toronto, and it is made accessible to everyone through the city's open data portal.

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- The instances that comprise the dataset represent Toronto residents and the various demographic and socio-economic characteristics they had at the time of the 2021 Census.
2. *How many instances are there in total (of each type, if appropriate)?*
	- There are over 2000 categories (e.g. education attainment, household size, etc.) of data provided for each of the 158 neighbourhoods in Toronto.
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- The short-form census is sent out to all households in a given region, and the long-form census is sent out to only 25% of households in a given region. With this, the 2021 Census data reflected in the Neighbourhood Profiles data can be a representative sample of all residents in a specific region.
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- Each instance consists of unprocessed text in an excel file. 
5. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- Missing data values can be imputed and the aggregation of individual household characteristics or experiences can lead to incorrect assumptions about certain regions and their populations.